Classical mechanics as the high-entropy limit \\ of quantum mechanics

### 2025-06-27 Articles/2024-ClassicalLimit/classicalLimitArticle.tex
For submission to Nature Physics
Fix bibliography capitalization and typo

\section{Introduction}

@@

One of the experimental challenges in quantum mechanics is producing states that are ``quantum enough'' to exhibit quantum properties. What we want to show that all these problems can ultimately be understood as reducing the entropy of the initial state.

@@

Mathematically, the trace distance $T(\rho,\sigma)$ between two mixed states $\rho$ and $\sigma$ is contractive under a completely positive trace-preserving (CPTP) map $M$. That is, $T(M(\rho),M(\sigma)) \leq T(\rho,\sigma)$~\cite{nielsen2010quantum}. In the finite-dimensional case, we can have a CPTP map that increases the entropy of all states except that of the maximally mixed state, which is a fixed point.\footnote{This is a particular instance of mixing/relaxing ergodic quantum channels where the fixed point is the maximally mixed state~\cite{Burgarth2012ErgodicAM}.}  Under such a map, all pure states are mapped to mixed states, orthogonal states are mapped to non-orthogonal states and the map is strictly contractive. In section 4, when we construct the high-entropy limit, the map that results will follow this intuition in the infinite-dimensional case, with the only difference that there is no maximally mixed state that does not change.\footnote{In the infinite-dimensional case, one could have infinite-dimensional subspaces that remain orthogonal while increasing the entropy of all its elements. This corresponds, for example, to the case where there are conserved quantities. Note that in the map we construct in section 4, the expectation of all polynomials of position and momentum changes, and therefore no function of position and momentum is a conserved quantity. }  If we call $\mathcal{G}$ the space of all mixtures of Gaussian states, the states with a non-negative Wigner function, given an arbitrary mixed state $\rho$, we will find a state $\sigma_{\rho} \in \mathcal{G}$ that is closest to $\rho$. As we apply our entropy-increasing map, the entropy of $\rho$ increases and $\sigma_{\rho}$ will get closer to $\rho$. In this sense the states will ``look classical.''

@@

>>>>

\section{Introduction}
\label{sec:intro}

@@

One of the experimental challenges in quantum mechanics is producing states that are ``quantum enough'' to exhibit quantum properties. What we want to show is that all these problems can ultimately be understood as reducing the entropy of the initial state.

@@

Mathematically, the trace distance $T(\rho,\sigma)$ between two mixed states $\rho$ and $\sigma$ is contractive under a completely positive trace-preserving (CPTP) map $M$. That is, $T(M(\rho),M(\sigma)) \leq T(\rho,\sigma)$~\cite{nielsen2010quantum}. In the finite-dimensional case, we can have a CPTP map that increases the entropy of all states except that of the maximally mixed state, which is a fixed point.\footnote{This is a particular instance of mixing/relaxing ergodic quantum channels where the fixed point is the maximally mixed state~\cite{Burgarth2012ErgodicAM}.}  Under such a map, all pure states are mapped to mixed states, orthogonal states are mapped to non-orthogonal states and the map is strictly contractive. In section~\ref{sec:high_entropy_limit}, when we construct the high-entropy limit, the map that results will follow this intuition in the infinite-dimensional case, with the only difference that there is no maximally mixed state that does not change.\footnote{In the infinite-dimensional case, one could have infinite-dimensional subspaces that remain orthogonal while increasing the entropy of all its elements. This corresponds, for example, to the case where there are conserved quantities. Note that in the map we construct in section 4, the expectation of all polynomials of position and momentum changes, and therefore no function of position and momentum is a conserved quantity. }  If we call $\mathcal{G}$ the space of all mixtures of Gaussian states, the states with a non-negative Wigner function, given an arbitrary mixed state $\rho$, we will find a state $\sigma_{\rho} \in \mathcal{G}$ that is closest to $\rho$. As we apply our entropy-increasing map, the entropy of $\rho$ increases and $\sigma_{\rho}$ will get closer to $\rho$. In this sense the states will ``look classical.''

@@

\section*{Acknowledgements}
This paper is part of the ongoing Assumptions of Physics project~\cite{aop-book}, which aims to identify a handful of physical principles from which the basic laws can be rigorously derived.

% TODO: include funding information
\paragraph{Funding information}
ML acknowledges funding via the FWF-funded QuantERA project QuSiED with project number I 6008-N.
